# Installation {#installation}

featureMSEA is available as a standalone R package and is also integrated into the **TidyMass Shiny** web application. Choose the option that best fits your workflow.

## Option A — TidyMass Shiny (recommended) {#shiny}

The easiest way to run featureMSEA is through the TidyMass Shiny interface, which provides a point-and-click workflow without any R coding.

**Access the online version:**

Visit [https://tidymassshiny.jaspershenlab.com](https://tidymassshiny.jaspershenlab.com){target="_blank"} — no installation needed.

**Run locally:**

```{r eval=FALSE}
# Install TidyMass and dependencies
remotes::install_gitlab("tidymass/tidymass")

# Install TidyMass Shiny
remotes::install_github("tidymass/tidymass_shiny")

# Launch
library(tidymass)
library(tidymassshiny)
run_tidymass_shiny()
```

Navigate to the **fMSEA Analysis** module under the **Tidymass Shiny Toolkits** dropdown.

## Option B — R package {#rpackage}

Install the `featureMSEA` R package directly for scripted workflows:

```{r eval=FALSE}
# Install from GitHub
remotes::install_github("tidymass/featureMSEA")

library(featureMSEA)
```

## System requirements {#requirements}

| Requirement | Minimum |
|-------------|---------|
| R version   | ≥ 4.1   |
| RAM         | 8 GB    |
| OS          | macOS, Windows, Linux |

> **Note for macOS users:** Cairo is required for graphics rendering. Install [XQuartz](https://www.xquartz.org){target="_blank"} first, then run `install.packages("Cairo", type = "source")`.

## Demo data {#demo-data}

Demo data for following the tutorial is available on [Google Drive](https://drive.google.com/file/d/1iuN0SvBybYlKzbHHSrqSRvKG-V18KG5Y/view?usp=drive_link){target="_blank"}.

The demo package contains:

- `feature_table.rda` — processed feature table
- `ms1_database.rda` — KEGG MS1 annotation database
- `pathway_database.rda` — KEGG pathway database
